Analysing the effects of COVID-19 on Crime in the United States

### 2022-04-30 outputs/paper/paper.Rmd
Added VAR modelling for each city, impulse response analysis+granger test and also fixed floating of figures.

output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib

@@

# Load LA data
# extract LA data from US data. 
covid19_LA <- covid19("USA", level = 3) |>
  filter(administrative_area_level_3 == "Los Angeles",
         administrative_area_level_2 == "California") |>
  # stayed at 1 for long time
  filter(confirmed > 1,
         date < "2020-06-12")

@@

Overall, our paper is a small contribution building upon and motivated by several recent papers that have covered the impact of the pandemic on the human psyche and crime rates across different states and countries [@boman2021global, @campedelli2020disentangling, @nivette2021global] alongside work prompting further research in the area [@eisner2020violence]. We achieve our goal of analyzing the pandemic data in conjunction with crime data obtained from the official city websites through time series modeling and also exploratory data analysis.

@@

```{r ,fig.cap = "Summary of several important variables present within the survey data",fig.height = 5, fig.width = 5, out.width = "300px", message = FALSE, echo = FALSE, warning=FALSE, fig.align="center"}
library(mapview)
Location <- c("Los Angeles","Chicago","Seattle","Philadelphia" )
lat <- c(33.82377,41.78798,47.714965, 39.952583)
lng <- c( -118.2668,-87.7738,-122.127166 ,-75.165222)
df <- data.frame(Location, lat,lng)

@@

```{r graph1, fig.cap = "Daily Confirmed Cases of COVID-19 in Chicago",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = chicago|>mutate(Date=as.Date(Date))

@@

```{r graph2, fig.cap = "Yearly Trends for the top five crimes in Chicago",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 3, fig.width = 10,}
# 2020 only
plt <- copy %>%
  dplyr::select(y_month, month, primary_type, year) %>%
  filter(primary_type %in% chicago_summary$primary_type[1:5] | primary_type == "MOTOR VEHICLE THEFT", y_month != "2020-06") %>%
  dplyr::count(year, month, primary_type) %>%
  na.omit() %>% 
  mutate(Date = as.Date(paste("2000", month, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = year, color = as.character(year))) + 
  geom_line() + 
  facet_wrap(~primary_type, nrow = 1) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

```{r graph3, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 6, fig.width = 12,}
# 2020 only

chicago_daily <- copy |>
  dplyr::select(Date, primary_type, year) |>
  filter(primary_type %in% chicago_summary$primary_type[1:5] | primary_type == "MOTOR VEHICLE THEFT", 
         year == 2020) |>
 dplyr::count(Date, primary_type) |>
  na.omit() |>
  ggplot(aes(Date, n, group = primary_type, color = primary_type)) +
  geom_line() +
  facet_wrap(~primary_type, nrow = 2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(chicago_summary$primary_type[1:5]))) +
  ylab('Number of Reported Crime') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
chicago_daily
# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))

@@

```{r graph5, fig.cap = "Daily Confirmed Cases of COVID-19 in L.A",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = la|>mutate(date_occ=as.Date(date_occ))

@@

# try first log difference
ts_diff_LA <- na.omit(diff(ts_LA))
plot.xts(ts_diff_LA,
         main = "Daily confirmed cases of COVID-19 in LA")
```

```{r graph6, fig.cap = "Yearly Trends for the top crimes in Los Angeles",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 3, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
ty <- copy %>%
  dplyr::select(y_month, month, crm_cd_desc, year) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5],
         y_month != "2020-06") %>%
  count(year, month, crm_cd_desc) %>%
  na.omit() %>%
  mutate(Date = as.Date(paste("2000", month, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n,
             group = year,
             color = as.character(year))) +
  geom_line() +
  facet_free(~crm_cd_desc) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

```{r graph7, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 6, fig.width = 12,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
daily <- copy %>%
  dplyr::select(date_occ, crm_cd_desc, year) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5],
         year == 2020) %>%
  count(date_occ, crm_cd_desc) %>%
  ggplot(aes(date_occ, n, group = crm_cd_desc, color = crm_cd_desc)) +
  geom_line() +
  facet_wrap(~crm_cd_desc, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(LA_summary$crm_cd_desc[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

daily
```

@@

```{r graph8, fig.cap = "Daily Confirmed Cases of COVID-19 in Seattle",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = seattle|>mutate(Date=as.Date(Date))

@@

```{r graph9, fig.cap = "Yearly Trends for the top crimes in Seattle",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 3, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
yty <- copy %>%
  dplyr::select(y_month, MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5],
         y_month != "2020-06") %>%
  dplyr::count(YEAR, MONTH, offense) %>%
  na.omit() %>%
  mutate(Date = as.Date(paste("2000", MONTH, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = YEAR, color = as.character(YEAR))) +
  geom_line() +
  facet_free(~offense, scales = 'free') +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b") +
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

```{r graph10, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 6, fig.width = 12,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")

seattle_daily <- copy %>%
  dplyr::select(Date, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR=='2020') %>%
  dplyr::count(Date, offense) %>%
  ggplot(aes(Date, n, group = offense, color = offense)) +
  geom_line() +
  facet_wrap(~offense, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(seattle_summary$offense[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

### Philadelphia

```{r graph11, fig.cap = "Daily Confirmed Cases of COVID-19 in Philadelphia",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = phil|>mutate(date=as.Date(date))

@@

```{r graph12, fig.cap = "Yearly Trends for the top crimes in Philadelphia",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 3, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
plt <- phil %>%
  dplyr::select(MONTH, dispatch_date, text_general_code, YEAR, y_month) %>%
  filter(text_general_code %in% phil_summary$text_general_code[1:5], y_month != "2020-06") %>%
  count(YEAR, MONTH, text_general_code) %>%
  na.omit() %>% 
  mutate(Date = as.Date(paste("2000", MONTH, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = YEAR, color = as.character(YEAR))) + 
  geom_line() + 
  facet_wrap(~text_general_code, nrow = 1) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))
plt
```

```{r graph13, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 6, fig.width = 12,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
daily <- copy %>%
  dplyr::select(date, text_general_code, YEAR, y_month) %>%
  filter(text_general_code %in% phil_summary$text_general_code[1:5], YEAR == 2020, y_month != '2020-06') %>%
  count(date, text_general_code) %>%
  na.omit() %>%
  ggplot(aes(date, n, group = text_general_code, color = text_general_code))+
  geom_line() + 
  facet_wrap(~text_general_code, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(phil_summary$text_general_code[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
daily

@@

\newpage
# Model 
# Results 
# Discussion
## First discussion point 
If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt from all
this.

## Second discussion point 
## Third discussion point 
## Weaknesses and next steps 
Weaknesses and next steps should also be included. 
\newpage 
<!-- \appendix  -->

>>>>

output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
urlcolor: blue

@@

# Load LA data
# extract LA data from US data. 
covid19_LA <- covid19("USA", level = 3) |>
  filter(administrative_area_level_3 == "Los Angeles",
         administrative_area_level_2 == "California") |>
  # stayed at 1 for long time
  filter(confirmed > 1)

@@

Overall, our paper is a small contribution building upon and motivated by several recent papers that have covered the impact of the pandemic on the human psyche and crime rates across different states and countries [@boman2021global, @campedelli2020disentangling, @nivette2021global] alongside work prompting further research in the area [@eisner2020violence]. We achieve our goal of analyzing the pandemic data in conjunction with crime data obtained from the official city websites, using R statistical programming [@citeR], through time series modeling and also exploratory data analysis.

@@

```{r ,fig.cap = "Geographical overview of the cities being examined under this study.",fig.height = 5, fig.width = 5, out.width = "300px", message = FALSE, echo = FALSE, warning=FALSE, fig.align="center"}
library(mapview)
Location <- c("Los Angeles","Chicago","Seattle","Philadelphia" )
lat <- c(33.82377,41.78798,47.714965, 39.952583)
lng <- c( -118.2668,-87.7738,-122.127166 ,-75.165222)
df <- data.frame(Location, lat,lng)

@@

## Chicago
The city of Chicago was chosen for this analysis because it has the highest crime rates in America. The gang culture and gun laws have made Chicago the hub of specifically violent crimes, with higher homicide rates on average as compared to other cities.  Furthermore, during the early stages of the COVID-19 pandemic, it was found that Chicago was one of the first few cities to harbor several virus strains [@anderson], leading to a disaster proclamation by then Governor Pritzker.  Chicago has since then seen roughly three waves of the virus as shown in Figure \@ref(fig:graph1). The data for the city of Chicago was sourced from the [Chicago police](https://home.chicagopolice.org/statistics-data/data-requests/), spanning six main categories of criminal activity which are assault, battery, criminal damage, deceptive practices, vehicle theft, and general theft.

```{r graph1, fig.cap = "Daily Confirmed Cases of COVID-19 in Chicago",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = chicago|>mutate(Date=as.Date(Date))

@@

# extract top 5 crime
top5crime <- chicago %>%
  filter(primary_type %in% head(chicago_summary$primary_type, 5)) %>%
  group_by(Date, primary_type) %>%
  tally() %>%
  spread(primary_type, n)

# rename columns
colnames(top5crime) <- c('time',
                         "assault",
                         "battery",
                         "criminal",
                         'deceptive',
                         "theft")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))

```
When closely examining yearly trends for the different crimes, Figure \@ref(fig:graph2), it can be observed that theft rates and battery noticeably went down in March of 2020 while crimes like assault and Vehicle Theft remained steady. Furthermore, it can be observed throughout the pandemic, these levels remained relatively consistent throughout 2021 and 2022. In general, it is also noted that most crimes, with the exception of vehicle theft and deception, went up generally in summer and then reduced when the temperatures fell for winter, indicating a seasonal trend.

```{r graph2, fig.cap = "Yearly Trends for the top five crimes in Chicago",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}
# 2020 only
plt <- copy %>%
  dplyr::select(y_month, month, primary_type, year) %>%
  filter(primary_type %in% chicago_summary$primary_type[1:5] | primary_type == "MOTOR VEHICLE THEFT", y_month != "2020-06") %>%
  dplyr::count(year, month, primary_type) %>%
  na.omit() %>% 
  mutate(Date = as.Date(paste("2000", month, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = year, color = as.character(year))) + 
  geom_line() + 
  facet_wrap(~primary_type, nrow = 1) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

Finally, a more granular look at the frequency of each crime only during the pandemic, as shown in Figure \@ref(fig:graph3), shows that most crimes during the pandemic remained relatively consistent with the exception of criminal damage and theft. A sudden uptick in criminal damage and theft is noted in June 2020, which can be attributed to the violence caused by certain bad actors during the George Floyd Protests.

```{r graph3, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 Onwards only

chicago_daily <- copy |>
  dplyr::select(Date, primary_type, year) |>
  filter(primary_type %in% chicago_summary$primary_type[1:5] | primary_type == "MOTOR VEHICLE THEFT", 
         year >= 2020) |>
 dplyr::count(Date, primary_type) |>
  na.omit() |>
  ggplot(aes(Date, n, group = primary_type, color = primary_type)) +
  geom_line() +
  facet_wrap(~primary_type, nrow = 2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(chicago_summary$primary_type[1:5]))) +
  ylab('Number of Reported Crime') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
chicago_daily
# ggplotly(chicago_?SZdaily, tooltip = c("Date", "n"))

@@

```{r graph5, fig.cap = "Daily Confirmed Cases of COVID-19 in L.A",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = la|>mutate(date_occ=as.Date(date_occ))

@@

# try first log difference
ts_diff_LA <- na.omit(diff(ts_LA))
plot.xts(ts_diff_LA,
         main = "Daily confirmed cases of COVID-19 in LA")

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui.Maecenas velit massa, fringilla vel tristique vitae, condimentum ut purus. Proin commodo eros iaculis, faucibus dui et, vehicula felis. Integer ligula urna, feugiat quis tristique vel, cursus nec quam. Mauris auctor sodales mauris, sed tristique quam sagittis in. Mauris ut bibendum enim. Maecenas consequat elit ut posuere aliquet. In hac habitasse platea dictumst. Integer est arcu, semper ac dolor at, pulvinar vehicula mi.

```{r graph6, fig.cap = "Yearly Trends for the top crimes in Los Angeles",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
ty <- copy %>%
  dplyr::select(y_month, month, crm_cd_desc, year) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5],
         y_month != "2020-06") %>%
  count(year, month, crm_cd_desc) %>%
  na.omit() %>%
  mutate(Date = as.Date(paste("2000", month, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n,
             group = year,
             color = as.character(year))) +
  geom_line() +
  facet_free(~crm_cd_desc) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

Nulla consectetur erat arcu, eu porta urna sagittis nec. In sed vestibulum turpis. Nullam vel libero eu sapien eleifend condimentum. Donec nunc mi, ornare quis viverra ut, volutpat vitae orci. Quisque sit amet porttitor mauris, faucibus maximus arcu. Duis et lorem eu elit consectetur posuere. Aliquam blandit nisl nisl. Proin velit arcu, gravida vel tellus.

```{r graph7, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
daily <- copy %>%
  dplyr::select(date_occ, crm_cd_desc, year) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5],
         year >= 2020) %>%
  count(date_occ, crm_cd_desc) %>%
  ggplot(aes(date_occ, n, group = crm_cd_desc, color = crm_cd_desc)) +
  geom_line() +
  facet_wrap(~crm_cd_desc, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(LA_summary$crm_cd_desc[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

daily
```
<!-- \newpage -->

@@

```{r graph8, fig.cap = "Daily Confirmed Cases of COVID-19 in Seattle",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}
# 2020 only
knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = seattle|>mutate(Date=as.Date(Date))

@@

```{r graph9, fig.cap = "Yearly Trends for the top crimes in Seattle",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
yty <- copy %>%
  dplyr::select(y_month, MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5],
         y_month != "2020-06") %>%
  dplyr::count(YEAR, MONTH, offense) %>%
  na.omit() %>%
  mutate(Date = as.Date(paste("2000", MONTH, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = YEAR, color = as.character(YEAR))) +
  geom_line() +
  facet_free(~offense, scales = 'free') +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b") +
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

Nulla consectetur erat arcu, eu porta urna sagittis nec. In sed vestibulum turpis. Nullam vel libero eu sapien eleifend condimentum. Donec nunc mi, ornare quis viverra ut, volutpat vitae orci. Quisque sit amet porttitor mauris, faucibus maximus arcu. Duis et lorem eu elit consectetur posuere. Aliquam blandit nisl nisl. Proin velit arcu, gravida vel tellus.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui.Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui.

```{r graph10, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")

seattle_daily <- copy %>%
  dplyr::select(Date, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR>='2020') %>%
  dplyr::count(Date, offense) %>%
  ggplot(aes(Date, n, group = offense, color = offense)) +
  geom_line() +
  facet_wrap(~offense, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(seattle_summary$offense[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))

@@

## Philadelphia

```{r graph11, fig.cap = "Daily Confirmed Cases of COVID-19 in Philadelphia",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 4, fig.width = 8,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
copy  = phil|>mutate(date=as.Date(date))

@@

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui. Maecenas velit massa, fringilla vel tristique vitae, condimentum ut purus. Proin commodo eros iaculis, faucibus dui et, vehicula felis. Integer ligula urna, feugiat quis tristique vel, cursus nec quam. Mauris auctor sodales mauris, sed tristique quam sagittis in. Mauris ut bibendum enim. Maecenas consequat elit ut posuere aliquet. In hac habitasse platea dictumst. Integer est arcu, semper ac dolor at, pulvinar vehicula mi.

```{r graph12, fig.cap = "Yearly Trends for the top crimes in Philadelphia",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 2, fig.width = 9,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
plt <- phil %>%
  dplyr::select(MONTH, dispatch_date, text_general_code, YEAR, y_month) %>%
  filter(text_general_code %in% phil_summary$text_general_code[1:5]) %>%
  count(YEAR, MONTH, text_general_code) %>%
  na.omit() %>% 
  mutate(Date = as.Date(paste("2000", MONTH, "01", sep = "-"))) %>%
  ggplot(aes(x=Date, y=n, group = YEAR, color = as.character(YEAR))) + 
  geom_line() + 
  facet_wrap(~text_general_code, nrow = 1) +
  guides(color = guide_legend(reverse = TRUE)) +
   ylab('') + xlab("")+
  theme(legend.title = element_blank()) +
  scale_x_date(date_labels = "%b")+
  theme(text = element_text(size=8), 
        plot.title = element_text(hjust = 0.5))
plt
```

Nulla consectetur erat arcu, eu porta urna sagittis nec. In sed vestibulum turpis. Nullam vel libero eu sapien eleifend condimentum. Donec nunc mi, ornare quis viverra ut, volutpat vitae orci. Quisque sit amet porttitor mauris, faucibus maximus arcu. Duis et lorem eu elit consectetur posuere. Aliquam blandit nisl nisl. Proin velit arcu, gravida vel tellus.

```{r graph13, fig.cap = "Frequency of each crime during months spanning the COVID-19 pandemic.",message = FALSE, echo = FALSE, warning=FALSE,fig.height = 5, fig.width = 10,}

knitr::opts_chunk$set(fig.pos = "!htbp", out.extra = "")
daily <- copy %>%
  dplyr::select(date, text_general_code, YEAR, y_month) %>%
  filter(text_general_code %in% phil_summary$text_general_code[1:5], YEAR >= 2020) %>%
  count(date, text_general_code) %>%
  na.omit() %>%
  ggplot(aes(date, n, group = text_general_code, color = text_general_code))+
  geom_line() + 
  facet_wrap(~text_general_code, nrow=2) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(phil_summary$text_general_code[1:5]))) +
  ylab('') + xlab("")+
  theme(legend.position = "none")+
  theme(text = element_text(size=10), 
        plot.title = element_text(hjust = 0.5))
daily

@@

\newpage
# Methods

## Vector autoregression Model (VAR)

Vector autoregression (VAR) is a statistical model used to capture the relationship between multiple quantities as they change over time. VAR is a type of stochastic process model. VAR models generalize the single-variable (univariate) autoregressive model by allowing for multivariate time series. VAR models are often used in economics and the natural sciences.
A pth-order VAR is denoted "VAR(p)" and sometimes called "a VAR with p lags". A pth-order VAR model is written as

$${\displaystyle y_{t}=c+A_{1}y_{t-1}+A_{2}y_{t-2}+\cdots +A_{p}y_{t-p}+e_{t},\,}y_{t}=c+A_{1}y_{{t-1}}+A_{2}y_{{t-2}}+\cdots +A_{p}y_{{t-p}}+e_{t},\,$$

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc.

## Granger Causality Test

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui. Maecenas velit massa, fringilla vel tristique vitae, condimentum ut purus. Proin commodo eros iaculis, faucibus dui et, vehicula felis. Integer ligula urna, feugiat quis tristique vel, cursus nec quam. Mauris auctor sodales mauris, sed tristique quam sagittis in. Mauris ut bibendum enim. Maecenas consequat elit ut posuere aliquet. In hac habitasse platea dictumst. Integer est arcu, semper ac dolor at, pulvinar vehicula mi.

<!-- \newpage -->

## Forecasting and Impulse Response

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec egestas justo in nulla bibendum ultrices. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec porttitor interdum purus, vel commodo lorem lacinia at. Suspendisse posuere ex non risus dapibus, vitae ullamcorper ex sollicitudin. Vestibulum interdum augue sit amet tortor mollis, vitae ornare diam pulvinar. Proin sapien elit, tempor non consectetur ut, mollis quis nunc. Maecenas tincidunt velit dolor, eu porta lacus ornare et. Ut in cursus felis. Etiam porta, ipsum sit amet pulvinar consectetur, nulla metus vehicula quam, ac blandit lacus ex et ante. Aliquam lacinia sagittis lobortis. Integer eros felis, gravida nec augue ut, blandit ultricies ipsum. Morbi aliquet at magna rhoncus vehicula. Suspendisse aliquam risus nunc, vitae eleifend lacus venenatis ut. Sed nisi eros, malesuada nec purus ut, tristique congue dui. Maecenas velit massa, fringilla vel tristique vitae, condimentum ut purus. Proin commodo eros iaculis, faucibus dui et, vehicula felis. Integer ligula urna, feugiat quis tristique vel, cursus nec quam. Mauris auctor sodales mauris, sed tristique quam sagittis in. Mauris ut bibendum enim. Maecenas consequat elit ut posuere aliquet. In hac habitasse platea dictumst. Integer est arcu, semper ac dolor at, pulvinar vehicula mi.

\newpage
# Results 
Nulla consectetur erat arcu, eu porta urna sagittis nec. In sed vestibulum turpis. Nullam vel libero eu sapien eleifend condimentum. Donec nunc mi, ornare quis viverra ut, volutpat vitae orci. Quisque sit amet porttitor mauris, faucibus maximus arcu. Duis et lorem eu elit consectetur posuere. Aliquam blandit nisl nisl. Proin velit arcu, gravida vel tellus.
